\hspace{-0.1cm}{J-ReCoVer}:~Java~Reducer~Commutativity~Verifier

### 2019-06-25 draft.tex
polish 1-4.1

MapReduce belongs among the most popular frameworks for data parallel
computation. A MapReduce program~\cite{dean04} consists of several pairs
of \emph{mappers} and \emph{reducers} running on a machine cluster for handling
big data in parallel. Usually, mappers and reducers are the only components in a
MapReduce program that involve concurrency. Mappers read data from a distributed
database and output a sequence of \emph{key-value} pairs. The elements of the
sequence (i.e., key-value pairs) with the same key are sent to the same reducer
for further processing. Due to scheduling policies and network latency, the same
inputs may arrive at a reducer in different orders in different executions.
Therefore, reducers are typically required to be \emph{commutative}, that is,
the output of a reducer is required to be independent of the order of its
inputs. The problem of checking whether this is indeed the case is known as the
\emph{commutativity problem} of
reducers~\cite{csallner13testing,xiao14mr,ChenHSW15,ChenSW16}. If all reducers
are commutative, they will have the same external behavior under all possible
schedules. In such a case, by fixing a schedule, the verification problem of a
MapReduce program can be reduced to the verification problem of a sequential
program, which is known to be much easier than the verification problem of
concurrent programs. One then suffices with considering any chosen interleaving
of input values instead of all of them.

@@

On the other hand, the non-commutative behavior of a reducer is often the source
of very tricky bugs. A study conducted by Microsoft investigated the
commutativity problem of $508$ reducers running on their MapReduce
server~\cite{xiao14mr}. These reducers were carefully checked using all
traditional means such as code review, testing, and experiments with real data
for months. Still, five of these programs contained very subtle bugs caused by
non-commutativity (which was confirmed by the programmers).

@@

In this paper, we present the J-ReCoVer tool (Java Reducer Commutativity
Verifier), which is available at~\verb|http://jrecover.iis.sinica.edu.tw|. The
tool implements a heuristic approach for checking the commutativity problem
that---despite its simplicity---works very efficiently on a large set of
practical integer reducer programs as shown by our experiments. The main
ingredient of the approach is a reduction from the commutativity problem to an
SMT problem. The reduction is incomplete but sound. It is accompanied with
several heuristics which enable the approach to scale to real-world examples.
For the case when the reducer is not proven commutative, we complement the
approach by using testing to find concrete counterexamples.

We collected benchmarks from open repositories such as Github and Bitbucket to
evaluate J-ReCoVer. With the help of a search engine \emph{searchcode.com} over
those repositories, we collected 118 programs. We provide this collection of
programs to other interested researchers as a side contribution of the paper.
Our tool J-ReCoVer is able to correctly analyze all but three of the programs.

@@

However, in our experience, if one naively reduces the commutativity problem to
an equivalence problem and checks it in a precise manner, many reducers cannot
be verified. Therefore, J-ReCoVer uses an over-approximation of the reducer's
behavior. This approximation allows for a much more efficient, yet---according
to our experiments---precise enough commutativity analysis.

@@

For presentation purposes, we provide a sketch of a highly simplified language
for describing reducers. The language allows us to give examples of reducers and
to discuss our approach for checking their commutativity. Let $\Var$ be a set of
\emph{integer variables}. An \emph{integer expression} in $\Exp$ can either be
a~variable from $\Var$, a constant value, a call to the $\cur$ function that
reads and consumes an input value of the reducer, their combinations over basic
arithmetic operations, or a non-deterministically chosen integer value $*$. A
\emph{command} in $\Cmd$ can be an assignment, a branch statement, a sequence of
commands, or an $out(v)$ statement that outputs the value of $v \in \Var$. A
\emph{reducer program} is defined as $s_1;\rloop\{s_2\};s_3$ where $s_1,s_2,s_3
\in \Cmd$. According to our observation over hundreds  of reducer programs in
open repositories, reducer programs are almost always in this form. The
$\rloop\{s_2\}$ statement enters the loop body to execute $s_2$ repeatedly until
the entire input list is consumed. To simplify presentation, we assume the reducer never output in the loop body $s_2$. An example of a reducer is shown in
Figure~\ref{fig:reducer_example}.

Some reducers use two (or more) top-level loops to compute the output, possibly
interleaved with some non-looping code. These loops are executed sequentially,
repeatedly iterating over the input list from its beginning. For example, for
calculating the standard deviation, one first computes the average of inputs and
then uses it to compute the final result (using two passes over the input list).
In that case, we suggest to verify the reducer by first partitioning it into two
(or more) reducers, each containing a single top-level loop, and then verifying
these reducers separately.\footnote{Nested loops can be removed by adding
additional branch statements since both the inner and outer loop are over the
same input list. In fact, such a program construct has never occurred in the
examples we have seen.} The top-level loops communicate through shared
variables. After the transformation, reducers corresponding to the second
top-level loop (and possibly further such loops) will work with random initial
values of the shared variable, which over-approximates the original behavior.
Usually, the second (or further) top-level loop are commutative even with
arbitrary initial shared variable values and hence J-ReCoVer can be used to
prove it.

@@

In order to be able to explain the transformation that we perform on an example,
consider the example in Figure~\ref{fig:reducer_max1}. The reducer presented in
the figure remembers the first input value in the variable $m$, increases its
value by 10, and then updates its value to bigger ones if any occur in the loop.
The main loop of the reducer is commutative in this case, but the reducer is not
commutative. A counterexample can be easily found. With the list $[1,2,3,4,5]$
and its permutation $[5,4,3,2,1]$ as the inputs, the reducer outputs $11$ and
$15$, respectively.

@@

To deal with the problem, we use a transformation that we first explain on the
example from Figure~\ref{fig:reducer_max1}. In particular, we move the prefix
$s_1$ into the loop body and use a new variable $s$ to force that the execution
of $s_1$ is always before the original loop body. The result after the
transformation is demonstrated in Figure~\ref{fig:reducer_max2}. The new reducer
program over-approximates all possible inputs/outputs of the original one. When
$s=1$, its output sequence is the same as for the original reducer. When $s\neq
1$, it may produce outputs that are not possible by the original reducer.
However, since the new reducer has all behaviors of the original one, if the new
reducer is commutative, the original one is also commutative.

@@

In general, the problem with the input before the loop can be handled as follows,
including the case where $\cur$ occurs multiple times in $s_1$. Assume that the
$s_1$ part of the reducer $s_1;\rloop\{s_2\};s_3$ has the form
$c_0;x_{1}{:=} \cur;c_1; \ldots; x_{m}{:=}\cur;c_m$, where $\cur$ does
not occur in $c_0,c_1,\ldots, c_m$. 
In the transformed reducer, the part before the loop is now $c_0;s:=1$. The loop body contains several branch statements. 
For all $j \in [1,m]$, we add a branch statement $\IF{s=j}{x_{j}:=\cur; c_j; s:=s+1}$. 
We also add $\IF{s=m+1}{s_2}$ for the original loop body.

@@

To handle the above issue, we perform a simple backward live variable analysis
to collect all variables whose value may propagate to the output command after
the loop execution. Only these variables are then required to be equivalent. In
our evaluation, the ratio of reducers that our approach can successfully analyze
is significantly increased---in particular, from $6.8~\%$ to $97.5~\%$---by
using this optimisation.

>>>>

MapReduce belongs among the most popular frameworks for data parallel
computation. A MapReduce program~\cite{dean04} consists of several pairs
of \emph{mappers} and \emph{reducers} running on a machine cluster for handling
big data in parallel. Usually, mappers and reducers are the only components in a
MapReduce program that involve concurrency. Mappers read data from a distributed
database and output a sequence of \emph{key-value} pairs. The elements of the
sequence (i.e., key-value pairs) with the same key are sent to the same reducer
for further processing. Due to scheduling policies and network latency, the same
inputs may arrive at a reducer in different orders in different executions.
Therefore, reducers are typically required to be \emph{commutative}, that is,
the output of a reducer is required to be independent of the order of its
inputs. The problem of checking whether this is indeed the case is known as the
\emph{commutativity problem} of
reducers~\cite{csallner13testing,xiao14mr,ChenHSW15,ChenSW16}. If all reducers
are commutative, they will have the same external behaviour under all possible
schedules. 
One then suffices with considering any chosen interleaving
of input values instead of all of them.
In such a case, by fixing a schedule, the verification problem of a
MapReduce program can be reduced to the verification problem of a sequential
program, which is known to be much easier than the verification problem of
concurrent~programs.

@@

On the other hand, the non-commutative behaviour of a reducer is often the source
of very tricky bugs. A study conducted by Microsoft investigated the
commutativity problem of $508$ reducers running on their MapReduce
server~\cite{xiao14mr}. These reducers were carefully checked using all
traditional means such as code review, testing, and experiments with real data
for months. Still, five of these programs contained very subtle bugs caused by
non-commutativity (which was confirmed by the programmers).

@@

In this paper, we present the J-ReCoVer tool (Java Reducer Commutativity
Verifier). A link to J-ReCoVer can be found in the supplementary materials. The
tool implements a heuristic approach for checking the commutativity problem
that---despite its simplicity---works very efficiently on a large set of
practical integer reducer programs as shown by our experiments. The main
ingredient of the approach is a reduction from the commutativity problem to an
SMT problem. The reduction is incomplete but sound. It is accompanied with
several heuristics which enable the approach to scale to real-world examples.
For the case when the reducer is not proven commutative, we complement the
approach by using testing to find concrete counterexamples.

We collected benchmarks from open repositories such as GitHub and Bitbucket to
evaluate J-ReCoVer. With the help of a search engine \emph{searchcode.com} over
those repositories, we collected 118 programs. We provide this collection of
programs to other interested researchers as a side contribution of the paper.
Our tool J-ReCoVer is able to correctly analyse all but three of the programs.

@@

However, in our experience, if one naively reduces the commutativity problem to
an equivalence problem and checks it in a precise manner, many reducers cannot
be verified. Therefore, J-ReCoVer uses an over-approximation of the reducer's
behaviour. This approximation allows for a much more efficient, yet---according
to our experiments---precise enough commutativity analysis.

@@

For presentation purposes, we provide a sketch of a highly simplified language
for describing reducers. The language allows us to give examples of reducers and
to discuss our approach for checking their commutativity. Let $\Var$ be a set of
\emph{integer variables}. An \emph{integer expression} in $\Exp$ can either be
a~variable from $\Var$, a constant value, a call to the $\cur$ function that
reads and consumes an input value of the reducer, their combinations over basic
arithmetic operations, or a non-deterministically chosen integer value $*$. A
\emph{command} in $\Cmd$ can be an assignment, a branch statement, a sequence of
commands, or an $out(v)$ statement that outputs the value of $v \in \Var$. A
\emph{reducer program} is defined as $s_1;\rloop\{s_2\};s_3$ where $s_1,s_2,s_3
\in \Cmd$. According to our observation over hundreds  of reducer programs in
open repositories, reducer programs are almost always in this form. The
$\rloop\{s_2\}$ statement enters the loop body to execute $s_2$ repeatedly for each input element until
the entire input list is consumed. An example of a reducer is shown in Figure~\ref{fig:reducer_example}. To simplify presentation, we assume in this paper the reducer never output in the loop body $s_2$. In the J-ReCoVer tool, an algorithm to deal with output inside the loop is implemented.

Some reducers use two (or more) top-level loops to compute the output, possibly
interleaved with some non-looping code. These loops are executed sequentially,
repeatedly iterating over the input list from its beginning. For example, for
calculating the standard deviation, one first computes the average of inputs and
then uses it to compute the final result (using two passes over the input list).
In that case, we suggest to verify the reducer by first partitioning it into two
(or more) reducers, each containing a single top-level loop, and then verifying
these reducers separately.\footnote{Nested loops can be removed by adding
additional branch statements since both the inner and outer loop are over the
same input list. In fact, such a program construct has never occurred in the
examples we have seen.} The top-level loops communicate through shared
variables. After the transformation, reducers corresponding to the second
top-level loop (and possibly further such loops) will work with random initial
values of the shared variable, which over-approximates the original behaviour.
Usually, the second (or further) top-level loop are commutative even with
arbitrary initial shared variable values and hence J-ReCoVer can be used to
prove it.

@@

In order to explain the problem with further details,
consider the example in Figure~\ref{fig:reducer_max1}. The reducer presented in
the figure remembers the first input value in the variable $m$, increases its
value by 10, and then updates its value to bigger ones if any occur in the loop.
The main loop of the reducer is commutative in this case, but the reducer is not
commutative. A counterexample can be easily found. With the list $[1,2,3,4,5]$
and its permutation $[5,4,3,2,1]$ as the inputs, the reducer outputs $11$ and
$15$, respectively.

@@

To deal with the problem, we apply a program transformation on the
example from Figure~\ref{fig:reducer_max1}. In particular, we move the prefix
$s_1$ into the loop body and use a new variable $s$ to force that the execution
of $s_1$ is always before the original loop body. The result after the
transformation is demonstrated in Figure~\ref{fig:reducer_max2}. The new reducer
program has the same inputs/outputs as the original one. Therefore, if the new reducer is commutative, the original one~is~also~commutative.

@@

In general, the problem with the multiple inputs before the loop can be handled as follows. Given a reducer $c_0;x_{1}{:=} \cur;c_1; \ldots; x_{m}{:=}\cur;c_m;\rloop\{s_2\};s_3$, where $m$ inputs are read before the loop and $\cur$ does
not occur in $c_0,c_1,\ldots, c_m$. 
In the transformed reducer, the part before the loop is now $c_0;s:=1$. The loop body contains several branch statements. 
For all $j \in [1,m]$, we add a branch statement $\IF{s=j}{x_{j}:=\cur; c_j; s:=s+1}$. 
We also add $\IF{s=m+1}{s_2}$ for the original loop body.

@@

To handle the above issue, we perform a simple backward live variable analysis
to collect all variables whose value may propagate to the output command after
the loop execution. Only these variables are then required to be equivalent. In
our evaluation, the ratio of reducers that our approach can successfully analyse
is significantly increased---in particular, from $6.8~\%$ to $97.5~\%$---by
using this optimisation.
